External Prior Guided Internal Prior Learning for Real Noisy Image Denoising

### 2016-11-08 CVPR2017_Review_Unified.tex
a

In the last decade, there are also several methods \cite{crosschannel2016,fullyblind,rabie2005robust,Liu2008,almapg,noiseclinic,ncwebsite,Zhu_2016_CVPR} developed for real noisy image denoising. Almost all these methods employ a two-stage framework: firstly estimating the parameters of the assumed noise model (usually Gaussian distribution or mixture of Gaussians (MoG)) and then performing denoising with the help of the estimated noise model. However, the noise in real noisy images are very complex and hard to be modeled by explict distributions. Therefore, the denoising performance of these methods would be influenced by the error between the estimated noise model and the actual noise distribution. This can be validated by the   denoised results listed in Figure \ref{fig1} ((g) and (h)), in which we evaluate the state-of-the-art real noisy image denoising method Noise Clinic \cite{noiseclinic,ncwebsite} and a commercial denoising software Neat Image \cite{neatimage}. From the results, we can see that the two methods do not perform well on the testing noisy image. These methods only exploit internal priors (e.g., noise prior, self-similarity) for real noisy image denoising and ignore the information of external natural clean images. 
\begin{figure*}
\centering
\subfigure{
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Noisy_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (a) Noisy \cite{crosschannel2016}: 33.30dB }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_CBM3D_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (b) BM3D \cite{bm3d,cbm3d}: 33.33dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_WNNM_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (c) WNNM \cite{wnnm}: 33.30dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_MLP_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (d) MLP \cite{mlp}: 34.22dB }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_CSF_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (e) CSF \cite{csf}: 35.39dB }
\end{minipage}
}\vspace{-3mm}
\subfigure{
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_TRD_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (f) TRD \cite{chen2015learning}: 35.97dB   }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_NC_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (g) NC \cite{noiseclinic}: 35.33dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_NI_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (h) NI \cite{neatimage}: 34.39dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Guided_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (i) Ours: \textbf{37.02}dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Mean_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (j) Mean Image \cite{crosschannel2016} }
\end{minipage}
}\vspace{-1mm}
\caption{Denoised images of the real noisy image ``Nikon D800 ISO 3200 A3" from \cite{crosschannel2016} by different methods. The images are better viewed by zooming in on screen.} 
\vspace{-3mm}
\label{fig1}
\end{figure*}

The external priors in natural clean images are free of the high correlation between noise and signals in real noisy images, while the internal prior is adaptive to the image and can recover better the latent clean image. Combining the priors of external clean images and priors of internalwould naturally achieves better performance. In this paper, we propose to employ the external priors learned from natural clean images to guide the learning of internal priors of given noisy image. We learn the external priors of patch groups \cite{pgpd} since nonlocal self-similarity is successfully used in many denoising methods \cite{nlm,bm3d,wnnm,pgpd}. The learning of external priors is performed by Gaussian Mixture Model. The learning of internal priors includes Gaussian selection and orthogonal dictionary learning (DL). The internal orthogonal DL process includes two alternating stages: updating sparse coefficients and updating orthogonal dictionary. The real noisy image denoising is simultaneously done with the DL process. Both of the two stages have closed-form solutions. Hence, our proposed denoising method is very efficient. Through comprehensive experiments on real noisy images captured by different cameras and settings, we demonstrate that the proposed method achieves better performance on real image denoising.

The rest of this paper will be summarized as follows: in Section 2, we briefly introduce the related work; in Section 3, we develop the proposed external prior guided internal prior learning model and formulate the overall image denoising algorithm; in Section 3, we demonstrate extensive experiments on real image denoising problem; in Section 4, we conclude our paper and give future work.

@@

In the last decade, researchers proposed many methods \cite{fullyblind,rabie2005robust,Liu2008,almapg,noiseclinic,ncwebsite,Zhu_2016_CVPR,crosschannel2016} for real image denoising problem. In \cite{fullyblind}, Portilla proposed to use scale mixture of Gaussians to estimate the noise covariance matrix and latent clean images in wavelet domain. The work of Rabie \cite{rabie2005robust} modeled the noisy pixels as outliers which are removed via Lorentzian robust estimator \cite{huber2011robust}. Liu \etal \cite{Liu2008} proposed to use ``noise level function" to estimate the noise and then use Gaussian conditional random field to obtain the latent clean image. Gong et al. \cite{almapg} m odels the noise by mixed $\ell_{1}$ and $\ell_{2}$ norms and remove the noise by sparsity prior in the wavelet transform domain. Later, Lebrun el al. proposed a multiscale denoising algorithm called ``Noise Clinic" \cite{noiseclinic,ncwebsite}. This method generalizes the NL-Bayes model \cite{nlbayes} to deal with blind noise and achieves state-of-the-art performance. Recently, Zhu et al. proposed a Bayesian model \cite{Zhu_2016_CVPR} which approximates and removes the noise via low-rank mixture of Gaussians. However, these methods largely depends on the modeling of noise in real noisy images which is hard to be modeled by explict distributions. Besides, the parametric estimation of the Gaussian or MoG distribution is often time consumming \cite{B}.

@@

In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. The overall objective log-likelihood function of GMM is
\begin{equation}\label{equ2}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation}
The learning process is similar to the learning stage of the PGPD method \cite{pgpd}. Please refer to \cite{pgpd} for more details. After the learning stage, we finally obtain a GMM model with $K$ Gaussian components. The learned parameters include mixture weights $\{\pi_{k}\}_{k=1}^{K}$, mean vectors $\{\boldsymbol{\upmu}_{k}\}_{k=1}^{K}$, and covariance matrices $\{\mathbf{\Sigma}_{k}\}_{k=1}^{K}$. Similar to \cite{pgpd}, the mean vector of each cluster is natural zeros, i.e., $\boldsymbol{\upmu}_{k}=\mathbf{0}$.

Now, we have clustered the PGs extracted from external clean images into $K$ Gaussians or subspaces. To better characterize each subspace, we perform singular value decomposition (SVD) on the covariance matrix:
\vspace{-2mm}
\begin{equation}\label{equ3}\vspace{-2mm}
\mathbf{\Sigma}_{k}=\mathbf{U}_{k}\mathbf{S}_{k}\mathbf{U}_{k}^{\top}.
\end{equation}
The singular vector matrices $\{\mathbf{U}_{k}\}_{k=1}^{K}$ are employed as the external orthogonal dictionary to guide the internal dictionary learning. The singular values in the diagonal of $\mathbf{S}_{k}$ reflect the significance of the singular vectors in $\mathbf{U}_{k}$ and utilized as prior weights for weighted sparse coding which will be discussed in next section.

@@

Given a real noisy image, assume we can totally extract $N$ local patches from it. Similar to the external prior learning stage, for the $n$-th local patch ($n=1,...,N$), we extract its $M$ most similar patches around it to form a noisy PG denoted by $\mathbf{Y}_{n} = \{\mathbf{y}_{n,1},...,\mathbf{y}_{n,M}\}$. Then the group mean of $\mathbf{Y}_{n}$, denoted by $\bm{\mu}_{n}$, is calculated and subtracted from each patch by $\mathbf{\overline{y}}_{n,m}=\mathbf{y}_{n,m}-\bm{\mu}_{n}$, leading to the mean subtracted PG $\mathbf{\overline{Y}}_{n}\triangleq \{\mathbf{\overline{y}}_{n,m}\}_{m=1}^{M}$. For adaptivity, we project the PG $\mathbf{\overline{Y}}_{n}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}_{n}$ is selected by firstly calculating the posterior probability of ``$\mathbf{\overline{Y}}_{n}$ belonging to the $k$-th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ4}\vspace{-1mm}
P(k|\mathbf{\overline{Y}}_{n})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{l})}
\end{equation}
for $k=1,...,K$, and then choosing the component with the maximum A-posteriori (MAP) probability $\max_{k}P(k|\mathbf{\overline{Y}}_{n})$.

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For the $k$-th subspace, assume the noisy PGs assigned to it are $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ such that $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$ and $\sum_{k=1}^{K}N_{k}=N$. We utilize the external orthogonal dictionary $\mathbf{U}_{k}$ (Eqn. (\ref{equ3})) to guide the learning of an orthogonal dictionary for adaptively characterizing the internal PGs in the $k$-th subspace. The learned dictionary defined as $\mathbf{D}_{k}\triangleq[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$ has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary $\mathbf{U}_{k}$, and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. For notation simplicity, we ignore the subspace index $k$ and denote the noisy PGs assigned to each subspace as $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning is performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-3mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
s.t.
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficient vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ in the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-1mm}
\begin{equation}\label{equ6}\vspace{-1mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer to Eqn. (\ref{equ3})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

>>>>

In the last decade, there are also several methods \cite{crosschannel2016,fullyblind,rabie2005robust,Liu2008,almapg,noiseclinic,ncwebsite,Zhu_2016_CVPR} developed for real noisy image denoising. Almost all these methods employ a two-stage framework: firstly estimating the parameters of the assumed noise model (usually Gaussian distribution or mixture of Gaussians (MoG)) and then performing denoising with the help of the estimated noise model. However, the noise in real noisy images are very complex and hard to be modeled by explict distributions. Therefore, the denoising performance of these methods would be influenced by the error between the estimated noise model and the actual noise distribution. This can be validated by the   denoised results listed in Figure \ref{fig1} ((g) and (h)), in which we evaluate the state-of-the-art real noisy image denoising method Noise Clinic \cite{noiseclinic,ncwebsite} and a commercial denoising software Neat Image \cite{neatimage}. From the results, we can see that the two methods do not perform well on the testing noisy image. These methods only exploit internal priors (e.g., noise prior, self-similarity) for real noisy image denoising and ignore the information of external natural clean images. 
\begin{figure*}
\centering
\subfigure{
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Noisy_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (a) Noisy \cite{crosschannel2016}: 33.30dB }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_CBM3D_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (b) BM3D \cite{bm3d,cbm3d}: 33.33dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_WNNM_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (c) WNNM \cite{wnnm}: 33.30dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_MLP_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (d) MLP \cite{mlp}: 34.22dB }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_CSF_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (e) CSF \cite{csf}: 35.39dB }
\end{minipage}
}\vspace{-3.5mm}
\subfigure{
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_TRD_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (f) TRD \cite{chen2015learning}: 35.97dB   }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_NC_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (g) NC \cite{noiseclinic}: 35.33dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_NI_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (h) NI \cite{neatimage}: 34.39dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Guided_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (i) Ours: \textbf{37.02}dB  }
\end{minipage}
\begin{minipage}[t]{0.195\textwidth}
\centering
\raisebox{-0.15cm}{\includegraphics[width=1\textwidth]{images/resize_br_Mean_CC_Noisy_Nikon_D800_ISO_3200_A3_66.png}}
{\footnotesize (j) Mean Image \cite{crosschannel2016} }
\end{minipage}
}\vspace{-1mm}
\caption{Denoised images of the real noisy image ``Nikon D800 ISO 3200 A3" from \cite{crosschannel2016} by different methods. The images are better viewed by zooming in on screen.} 
\vspace{-3mm}
\label{fig1}
\end{figure*}

The external priors in natural clean images are free of the high correlation between noise and signals in real noisy images, while the internal prior is adaptive to the image and can recover better the latent clean image. Combining the priors of external clean images and priors of internalwould naturally achieves better performance. Based on above considerations, in this paper, we propose to employ the external priors learned from natural clean images to guide the learning of internal priors of given noisy image.\ The flowchart of the proposed method is illustrated in Fig.\ \ref{fig2}.\ In the external stage, we learn the external priors of patch groups \cite{pgpd} since nonlocal self-similarity is successfully used in many denoising methods \cite{nlm,bm3d,wnnm,pgpd}.\ The learning of external priors is performed by Gaussian Mixture Model. In the internal stage, the learning of internal priors includes Gaussian selection and orthogonal dictionary learning (DL).\ The internal DL process includes two alternating stages: updating sparse coefficients and updating orthogonal dictionary. The image denoising is simultaneously done with the DL process.\ Both of the two stages have closed-form solutions.\ Hence, our proposed denoising method is very efficient.\ Through extensive experiments on real noisy image denoising, we demonstrate that the proposed method achieves better performance on real image denoising.

The rest of this paper will be summarized as follows: in Section 2, we briefly introduce the related work; in Section 3, we develop the proposed external prior guided internal prior learning model and formulate the overall image denoising algorithm; in Section 3, we perform extensive experiments on real image denoising problem; in Section 4, we conclude our paper and give future work.

@@

In the last decade, researchers proposed many methods \cite{fullyblind,rabie2005robust,Liu2008,almapg,noiseclinic,ncwebsite,Zhu_2016_CVPR,crosschannel2016} for real image denoising problem. In \cite{fullyblind}, Portilla proposed to use scale mixture of Gaussians to estimate the noise covariance matrix and latent clean images in wavelet domain. The work of Rabie \cite{rabie2005robust} modeled the noisy pixels as outliers which are removed via Lorentzian robust estimator \cite{huber2011robust}. Liu \etal \cite{Liu2008} proposed to use ``noise level function" to estimate the noise and then use Gaussian conditional random field to obtain the latent clean image. Gong et al. \cite{almapg} m odels the noise by mixed $\ell_{1}$ and $\ell_{2}$ norms and remove the noise by sparsity prior in the wavelet transform domain. Later, Lebrun el al. proposed a multiscale denoising algorithm called ``Noise Clinic" \cite{noiseclinic,ncwebsite}. This method generalizes the NL-Bayes model \cite{nlbayes} to deal with blind noise and achieves state-of-the-art performance. Recently, Zhu et al. proposed a Bayesian model \cite{Zhu_2016_CVPR} which approximates and removes the noise via low-rank mixture of Gaussians. However, these methods largely depends on the modeling of noise in real noisy images which is hard to be modeled by explict distributions. Besides, the parametric estimation of the Gaussian or MoG distribution is often time consumming.

@@

In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. The overall objective log-likelihood function of GMM is
\begin{equation}\label{equ1}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation}
The learning process is similar to the learning stage of the PGPD method \cite{pgpd}. Please refer to \cite{pgpd} for more details. After the learning stage, we finally obtain a GMM model with $K$ Gaussian components. The learned parameters include mixture weights $\{\pi_{k}\}_{k=1}^{K}$, mean vectors $\{\boldsymbol{\upmu}_{k}\}_{k=1}^{K}$, and covariance matrices $\{\mathbf{\Sigma}_{k}\}_{k=1}^{K}$. Similar to \cite{pgpd}, the mean vector of each cluster is natural zeros, i.e., $\boldsymbol{\upmu}_{k}=\mathbf{0}$.

Now, we have clustered the PGs extracted from external clean images into $K$ Gaussians or subspaces. To better characterize each subspace, we perform singular value decomposition (SVD) on the covariance matrix:
\vspace{-2mm}
\begin{equation}\label{equ2}\vspace{-2mm}
\mathbf{\Sigma}_{k}=\mathbf{U}_{k}\mathbf{S}_{k}\mathbf{U}_{k}^{\top}.
\end{equation}
The singular vector matrices $\{\mathbf{U}_{k}\}_{k=1}^{K}$ are employed as the external orthogonal dictionary to guide the internal dictionary learning. In Figure \ref{fig3} (a) and (b), we illustrate an external clean image and one orthogonal dictionary learned via GMM on PGs of the external clean image. To better illustrate the dictionary The singular values in the diagonal of $\mathbf{S}_{k}$ reflect the significance of the singular vectors in $\mathbf{U}_{k}$ and utilized as prior weights for weighted sparse coding which will be discussed in next section.

@@

Given a real noisy image, assume we can totally extract $N$ local patches from it. Similar to the external prior learning stage, for the $n$-th local patch ($n=1,...,N$), we extract its $M$ most similar patches around it to form a noisy PG denoted by $\mathbf{Y}_{n} = \{\mathbf{y}_{n,1},...,\mathbf{y}_{n,M}\}$. Then the group mean of $\mathbf{Y}_{n}$, denoted by $\bm{\mu}_{n}$, is calculated and subtracted from each patch by $\mathbf{\overline{y}}_{n,m}=\mathbf{y}_{n,m}-\bm{\mu}_{n}$, leading to the mean subtracted PG $\mathbf{\overline{Y}}_{n}\triangleq \{\mathbf{\overline{y}}_{n,m}\}_{m=1}^{M}$. For adaptivity, we project the PG $\mathbf{\overline{Y}}_{n}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}_{n}$ is selected by firstly calculating the posterior probability of ``$\mathbf{\overline{Y}}_{n}$ belonging to the $k$-th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ3}\vspace{-1mm}
P(k|\mathbf{\overline{Y}}_{n})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{l})}
\end{equation}
for $k=1,...,K$, and then choosing the component with the maximum A-posteriori (MAP) probability $\max_{k}P(k|\mathbf{\overline{Y}}_{n})$.

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For the $k$-th subspace, assume the noisy PGs assigned to it are $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ such that $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$ and $\sum_{k=1}^{K}N_{k}=N$. We utilize the external orthogonal dictionary $\mathbf{U}_{k}$ (Eqn.\ (\ref{equ2})) to guide the learning of an orthogonal dictionary for adaptively characterizing the internal PGs in the $k$-th subspace. The reasons we aim to learn orthogonal dictionaries are two-fold: firstly, given an orthonormal dictionary  $\mathbf{D}$, its quality measure \emph{mutual incoherence} $\mu(\mathbf{D})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}}$
is naturally $0$ and therefore better than other redundant dictionaries; secondly, the orthogonality of dictioanry can guarantee the dictionary learning stage has closed-form solutions (please refer to Eqn.\ (\ref{equ9})) and the closed-form solutions will make our proposed method efficient, which will be discussed in experimental section.

The learned dictionary $\mathbf{D}_{k}\triangleq[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$ has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary $\mathbf{U}_{k}$, and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. For notation simplicity, we ignore the subspace index $k$ and denote the noisy PGs assigned to each subspace as $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning is performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ4}\vspace{-3mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
s.t.
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficient vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ in the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-1mm}
\begin{equation}\label{equ5}\vspace{-1mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer to Eqn. (\ref{equ2})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}
